---
title: "Untitled"
author: "Savannah Weaver"
date: "2023-04-06"
output: html_document
---



# Packages


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```



```{r}

```

Notes from Tues 4/11 lecture:

- cntrl+option+i = insert code chunk
- cntrl+shift+m = insert pipe
- cntrl+shift+1/2/3/4 to zoom in and out of different panes
- naming chunks is helpful bc you can then click on them, like the outline 
- janitor pkg for cleaning up column names etc
- read_csv loads a tibble, while read.csv loads as a dataframe



# Part 1: QQ Plot Function

https://www.youtube.com/watch?v=n4-ZiJq_nSA


Background: Q-Q Plots
A Q-Q (Quantile-Quantile) Plot is a way of checking if two collections of observations come from the same distribution.

The steps are as follows:

Take two vectors of the same length, x and y.

Put the vectors in order from largest to smallest.

Pair the ordered vectors up, so that the smallest value of x is paired with the smallest value of y, and so on.

Make a scatterplot of the ordered pairs.

If the points in the scatterplot fall on a straight line, with intercept 0 and slope 1, this suggests that x and y are sampled from the same distribution.

In this section, you will use a Q-Q Plot to check if a vector of values x comes from a Normal distribution.
The approach is to randomly generate a new vector y from a Normal distribution with the same mean and standard deviation as x, then to create a Q-Q plot of x and y. (Note: A typical Normal Q-Q plot uses theoretical quantiles instead of randomly generated values. We're taking a bit of a shortcut in this assignment.)

Write a function (100 points)
Your final function should take as input a numeric vector.

It should return (not just print!) a Q-Q Plot comparing your input to Normally distributed values.

You may not use any existing functions specific to Q-Q plots; including (but not limited to) qqplot(), geom_qq, or stat_qq.

Demonstrate that your function works by running it on either real data of your choice, or on a non-Normal vector that you create.



create a normal distribution of length n:

```{r}
rnorm(10, # n
      50 # mean around which dist is set
      # can also do SD, but we will leave as default
      )
```




also have a random list of (normally I guess) distributed values:

```{r}
example_dist <- rnorm(20, 30)
example_dist
sort(example_dist)
```


plot:

```{r}
ggplot() + 
  geom_point(aes(x = example_dist, # make this the normal distribution
                 y = example_dist # make this the comparison one
                 )) +
  theme_classic() +
  # add labels
  xlab("Normal Distribution") +
  ylab("Your Data") +
  ggtitle("QQ Plot")
```


load a comparison dataset:

```{r}
iris
```



## Put it all into the function

This function takes a vector of numbers, compares it to a normal distribution, then returns a comparison plot, AKA a Q:Q plot. One assumption is that it expects a vector (vs a dataframe or tibble).


*notes*:

- instead of writing the function in a code chunk, it might be better to write it in a separate R file
- then, use function source("R script filename") to basically run everything in that script, to give you the function saved there!

```{r}
QQ_plot_my_func <- function(dist_compare) {
  
  # first create a comparison normal distribution
  # NOTE: could have saved the values I pass to rnorm beforehand,
  #   that way the functions are not nested within the rnorm function
  reference_norm <- rnorm(length(dist_compare), # number of points needed
                          mean(dist_compare), # mean around which dist is set
                          sd(dist_compare) # sd around which dist is set 
                                           # (this centers it on the 1:1 comparison line)
                          )
  
  # order both least to greatest & pair
  comp_df <- tibble(refs_ordered = sort(reference_norm),
                    yours_ordered = sort(dist_compare))
  
  # then plot them
  ggplot(comp_df) + 
    geom_point(aes(x = refs_ordered, # make this the normal distribution
                   y = yours_ordered # make this the comparison one
                   )) +
    theme_classic() +
    # add labels
    xlab("Normal Distribution") +
    ylab("Your Data") +
    ggtitle("QQ Plot") +
    # add a reference line
    geom_abline(slope = 1, intercept = 0,
                linetype = "dashed", alpha = 0.5) -> plot
  
  plot # don't need to explicitly state "return"
  # for ex, return(plot) # might be easier to SEE "return" when looking at function
  
}
```

corrections to my function:

- 





try it:

```{r}
QQ_plot_my_func(iris$Sepal.Width)
```






# Part 2


Bootstrapping is the only part of this example I do not know well, but I need to review first.

Oftentimes, coding starts at the end: Where do I want to end up, and how do I get there?



## BOOTSTRAPPING

https://www.youtube.com/watch?v=IPQaqi6rP0o

function notes

- replicate() repeats what you tell it to, over and over
- 
```{r}
# usually, create a function
# input can be a vector or a df
my_boot_mean <- function(vect) {
  
  # replicate sampling of the dataset
  replicate(100, # number of obs in a resample (NOT how many samples)
            sample(dat$col, nrow(dat), replace = TRUE) # what we want to resample from
            ) -> reps
  
  apply(reps, 2, # 2 = ?
        funct to do across bootreps)
}



# tidy_boots is a tibble OF the resamples = VERY nested structure
# literally creates a CLASS called resample that remembers what df it came from
tidy_boots <- bootstrap(df, # give it the whole dataframe
                        100 # number of times it is resampled
                        )

# but how do I decide how BIG a resample is?

# then aply
map(tidy_boots,
    function(x),
    func(df$col)
    )
```





try bootstrapping on the iris dataset to get the mean Sepal.Length

```{r}
# dataset
head(iris)

# pkg for tidyverse version of bootstrapping
library(tidymodels)

# create resamples
iris_boots <- bootstraps(iris, # dataframe
                         times = 10 #number of samples
                         )

# create a function for a SLR of Sepal.Length ~ Sepal.Width
my_lm <- function(resamp_df) {
  
  lm <- lm(data = resamp_df, Sepal.Length ~ Sepal.Width)
  
  slopes <- coefficients(lm)[2]
  
  slopes
}

# then apply function to each sample
boot_models <-
  iris_boots %>% 
  mutate(model_slope = map(splits, my_lm))

# then have to somehow get mean & CI
# this does not work
# mean(boot_models$model_slope)
```











