---
title: "random notes"
author: "Savannah Weaver"
date: "2023-04-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(leaflet)
library(gganimate)
library(plotly)
library(ggmap)
```








## rmd pro

in the code chunk name ({r}), if you add cache=TRUE, then it will save things, and reuse it unless you change the code that made it - **BUT** if you edit things elsewhere in the doc and not in the cached chunk, then the cache chunk won't know to update... a source of hidden errors and heartbreak 




# Interactive Mapping!!!


leaflet and plotly use this thing called reactjs, which allows images to react to your mouse - VS - Shiny apps use R code to allow things to be interactive to the user


## gganimate 

gganimate is nice bc you can just add it to an existing ggplot
AND works anywhere that you could put a gif (like scientific presentations!)
also need to be able to make gifs on your computer (i.e. gifski pkg)
usually use "transition_states()"
then, after adding that geom, you run the plot object through the animate() function, which also has some settings that can be made:

animate(
plot_object_name,
nframes = n
)


## plotly

this gives an object type of **html widget** - not something that works to save as an image that can go in Word or PDF... BUT, the pro is that it can be interactive - for example, in Shiny! because it's super easy to hover over things to see more info :)


```{r}
iris %>%
  plot_ly(x = ~Species,
          y = ~Sepal.Length,
          color = ~Species,
          type = "box") %>% # note that we PIPE bc add new function, not adding gg layer
  # add extra things for hoverng over
  add_trace(x = ~Species, # for what 
            mode = "markers", # how get popup
            name = ~Species # what to pop up
            )
```






## leaflet

oftentimes, if we want interactive things, we specifically want it for a **MAP**!

leaflet structure is more similar to ggplot than plotly, but still use pipes instead of pluses

```{r}
iris_map <- iris %>%
  mutate(randomize = rnorm(nrow(.)),
         lats = 35.2828 + randomize,
         longs = -120.6596 + randomize
         ) 

leaflet() %>%
  addTiles() %>%
  addMarkers(lat = iris_map$lats, lng = iris_map$longs, popup = iris_map$Species)

#^LOL




leaflet() %>%
  # automatically get all the boundaries of the world continents, countries, etc
  addTiles() %>%
  # can add pins/markers with lat and lng
  addMarkers(lat = 35.2828, lng = -120.6596, 
             popup = "Madonna" # popup when you click on it
             )
```